---
title: "Post-Impute_RSV"
author: "Mari Johnson"
date: "05/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = '~/RSV/data/post-imp/Assoc')
```
```{r Packages}
  library(dplyr)
  library(data.table)
  library(ggplot2)
  library(qqman)
  library(ggrepel) 
  library(tidyr)
  library(tidylog)
  library(stringi)
  library(janitor)
  library(stringr)
  library(RColorBrewer)
```


### Concatenation

Concatenate separate chromosome files from imputation server back into single merged file. Script utilises bcftools.

```{bash}
# Call concatenation script
bash concat.sh
```

### QC imputed SNPs

Re-check previous thresholds

```{bash PLINK set up and QC}

#Convert VCF to plink2 pfile

plink2 --vcf RSV_imputed.vcf.gz --geno 0.1 --mind 0.1 --make-pgen --out RSV_imp_QC
plink2 --pfile RSV_imp_QC --geno 0.05 --mind 0.05 --maf 0.01 --make-pgen --out RSV_imp_QC
#SNPs missing from 5% of population MAF <5 %
#So out of 340 the SNPs would have to be in 12 people
#10% cut off is probably better for assoc analysis in 34 people

#Convert back to bfile
#Plink2 pmerge function is still under development

plink2 --pfile RSV_imp_QC --geno 0.05 --mind 0.05 --maf 0.01 --make-bed --out RSV_imp_QC

```

**RSV post-imputation results:**

-   No variants removed after first QC round

-   25531828 variants removed due to allele frequency threshold(s)\
    8,917,131 variants remaining after main filters

```{bash Run Ethnicity Script}
# Script to check ethnicity data against 1000 Genomes
bash ancestry_check_38.sh

```

```{bash Filter IDs}
# Just keep our case/control study samples for now
plink2 --pfile RSV_imp_QC --keep study_samples.txt --make-pgen --out RSV_Oxford
# 318 samples

# Convert plink 1 version
plink2 --pfile RSV_Oxford --make-bed --out RSV_Oxford
```

### PCA

-   Use a PCA to account for population structure in data

-   Check ancestry first by plotting PCA and colour coding by ethnicity

-   Can also merge with public population databases

-   Include PCA's as covariates in association model

-   Need to prune for LD and sibling pairs (IBD)


```{bash sib check}

# Requires 1.9
plink --bfile RSV_Oxford --genome --missing --out RSV_Oxford_sib

#Remove one sample from each pair with pi-hat (% IBD) above threshold (0.1875 below):
awk '{if ($10 >= 0.1875) {print $1, $2}}' RSV_Oxford_sib.genome | uniq > RSV_Oxford_sib.outliers.txt 
awk '{if ($10 >= 0.1875) {print $1, $2, $3, $4, $10}}' RSV_Oxford_sib.genome | uniq > RSV_Oxford_sib.report.txt 
wc -l RSV_Oxford_sib.outliers.txt

# Prune samples from study when calculating PCD
# Check meta data too
```

```{bash}
awk '{if ($10 >= 0.1875) {print $1, $2}}' RSV_Oxford_sib.genome | uniq > RSV_Oxford_sib.outliers.txt

#Filter in plink
plink2 \
--bfile RSV_Oxford \
--remove RSV_Oxford_sib.outliers.txt \
--make-bed \
--out RSV_Oxford_nosib
```

```{bash Run PCA}
# PCA prep LD prune and IBD check

plink2 --bfile RSV_Oxford_nosib --pca 


#plink2 --pfile RSV_Oxford --pca 
```

Run PCA in R
There is also the PCA plotting script from the ancestry analysis

```{r Set up PCA in R, eval=TRUE}
pca <- fread("plink2.eigenvec")
eigenval <- fread("plink2.eigenval")
#Remove hashtag manually in IID in txt file, or next time try '#'

covar <- fread("covar.txt") #"covar_all.txt" 
pca_covar <- left_join(pca, covar, by = "IID") 

#matched 318 rows :)

pca_covar$Sex <- as.factor(pca_covar$Sex)
levels(pca_covar$Sex) <- c("Male", "Female")

#ooh could compare new pcs.y with old pcs.x
pca_covar %>% ggplot(aes(PC2.x, PC2.y)) + geom_point()
# A couple of outliers but no major difference



# Plot PCA coloured by Sex
pca1 <- pca_covar %>%
  ggplot(aes(PC1.y, PC2.y, color = Sex)) +
  geom_point() +
  labs(x = "PC1", y = "PC2", title = "PCA by Sex")
pca1

# Rearrange table
pca_covar <- pca_covar[,c(1:2,13:16,7:12,17:19)]
names(pca_covar) <- str_remove_all(names(pca_covar), ".y")


# Calculate Percentage variance explained
pve <-
  data.frame(PC = 1:10, pve = round(eigenval / sum(eigenval) * 100, 2))

colnames(pve) <- c("PC", "pve")
pve$PC <- as.factor(pve$PC)

p <- pve %>% ggplot(aes(x = PC, y = pve))

pve %>% ggplot(aes(x = PC, y = pve, fill = PC)) +
  geom_col() +
  geom_line() +
  labs(x = "Principle Component",
       y = "Percentage Varience Explained")

#Elbow at PC4 #Use first 4 PCAs as Covars

# Make new covar file to use
write.table(pca_covar, file = "pca_covar2.txt", row.names = F, quote = F, sep = "\t")

```

### Conclusions

-   Include first 4 PC's as covariates in association analysis

### Association Analysis

-   Initial linear association model (GLM) with --pheno-name RESV_total_score

-   Covariates to include = PC1-4, Sex, has_pre_ex_con, baseline_age_at_visit\
    <https://www.cog-genomics.org/plink/2.0/assoc>

```{bash Association Analysis}
#mkdir -p Assoc
plink2 --pfile RSV_Oxford --pheno pheno.txt --pheno-name RESV_total_score --covar covar.txt --covar-name Sex, PC1, PC2, PC3, PC4, has_pre_ex_con, baseline_age_at_visit --covar-variance-standardize --glm --out Assoc/
```
```{bash Case/Control Analysis}
#mkdir -p Assoc
plink2 --pfile RSV_Oxford --pheno pheno.txt --pheno-name RSV --covar covar.txt --covar-name Sex, PC1, PC2, PC3, PC4, has_pre_ex_con, baseline_age_at_visit --covar-variance-standardize --glm --out Assoc/
```

```{r Plot Assoc Results, eval = TRUE}

gwas <- fread("Assoc/RESV_total_score.glm.linear")

#rename cols for qqman
colnames(gwas) <- c("CHR", "BP", "SNP", "REF", "ALT", "A1","TEST","OBS_CT","BETA",        
                           "SE","T_STAT","P","ERRCODE")

gwas$TEST <- as.factor(gwas$TEST)
gwas<- filter(gwas, TEST == "ADD") #Keep just genetic testing

#Plot Manhattan
library(qqman)
manhattan(gwas, ylim = c(0, 8), cex = 0.6, cex.axis = 0.9, genomewideline = F, col = c("mediumseagreen", "darkolivegreen4"))

#Inspect top snps:
tophits <- filter(gwas, P < 10e-5)
head(tophits)
#Write top hits file 
write.table(tophits, file = "Assoc/tophits.txt", quote = F, row.names= F, sep = "\t")
```

```{bash CD14 association}
plink2 --pfile RSV_Oxford --chr 5 --from-bp 140630856 --to-bp 140634071 --pheno pheno.txt --pheno-name RESV_total_score --covar covar.txt --covar-name Sex, PC1, PC2, PC3, PC4, has_pre_ex_con, baseline_age_at_visit --covar-variance-standardize --glm --out CD14_
```

```{r CD14 results}

```
